---
title: "ML Project"
author: "Hector Munoz"
date: "24 de octubre de 2017"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data import and preparation
In the following section the data was prepared for the machine learning process. The 20 observations for the final quizz evaluation were stored in a dataframe labeled "final_test", while the rest of  the data was stored in the variable "training". Finally, columns where all entries were missing values (NA) were removed from the "trainig" DF. 

```{r}
suppressPackageStartupMessages(suppressWarnings(library(caret))) 
suppressPackageStartupMessages(suppressWarnings(library(data.table))) 
suppressPackageStartupMessages(suppressWarnings(library(dtplyr))) 
suppressPackageStartupMessages(suppressWarnings(library(dplyr))) 
suppressPackageStartupMessages(suppressWarnings(library(tidyr))) 

```


```{r import, cache = TRUE}
# Data import and setup ---------------------------------------------------

#Download of training file
    #download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml_training.csv","pml_training") 
    training <- read.csv("C:\\Users\\Hp\\Documents\\pml_training",
                         na.strings = c("","NA","#DIV/0!"))
#Download of testing file
    #download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml_testing.csv","pml_test")
    final_test <- read.csv("C:\\Users\\Hp\\Documents\\pml_test",
        na.strings = c("","NA","#DIV/0!"))
    #NOTE: 

# Data cleaning -----------------------------------------------------------
#Remove unnecessary columns
    # Convert to data.table
    training <- training %>% tbl_dt()
    training_raw <- copy(training)
    training <- copy(training_raw)
    final_test <- final_test %>% tbl_dt()
    
    # Inspect missing values
    #sapply(training, function(x) mean(!is.na(x))) %>% sort()
    
    #Remove counter and "all-zero" column
    training <- training %>% 
        dplyr::select(-X, -kurtosis_yaw_belt, -cvtd_timestamp,
                      -skewness_yaw_belt, -kurtosis_yaw_dumbbell,
                        -skewness_yaw_dumbbell, -kurtosis_yaw_forearm, -skewness_yaw_forearm)
    

```

## Data partition
A test set with 20% of the observations (disting from the 20-observation "final_test" dataframe) was built by taking 3923 rows at random out of the training set. The new "training" variable now contains therefore, 3923 less observations. 
```{r}
# Subdivide training set ------------------------------------------------
    # Create data partition in order to estimate out-of-sample error
    index <- createDataPartition(training$classe, p = 0.8, list = F)
    test <- training[-index]
    training <- training[index]
```

## Model building and assessment
The training set was preprocessed through the removal of the zero-variance ("zv") and nea-zero-variance ("nzv") columns. After that, the median value for each column was imputed in rows where values were missing ("medianImpute""). Finally, the principal components were obtained from the dataset ("pca").  

The ML model was built using a *randomforest*. The cross-validation method selected was a 3-fold CV, as indicated in train_control. 
```{r cache = TRUE}
# Model -------------------------------------------------------------------
    # PreProcess the data
    preProcess_model <- preProcess(x = training, 
                                   method = c("zv","nzv","medianImpute", "pca"))
    prep_training <- predict(preProcess_model, training)    
    browser()
    # Apply Cross-Validation
    train_control <- trainControl(method = "cv", number = 3)
#Initial basic model
    model <- train(classe~.,data = prep_training,method="rf", 
                   trainControl = train_control)
    
# Model assessment
(train_asessment <- confusionMatrix(training$classe, predict(model, prep_training)))

prep_test <- predict(preProcess_model, test)
(test_asessment <- confusionMatrix(test$classe, predict(model, prep_test)))
```

In this case, the training set accuracy and the test set accurracy (i.e. the out-of sample accuracy estimate) were calculated in the tables above. 